\scshape Evolutionary dynamics, equilibrium selection, and what population and quantitative genetics have taught evolutionary game theory

### 2022-10-10 manuscript.tex
moar writing

Notably, the conditions in \eqref{eq:ess:def} do not reference the genetic basis of the trait $X$ nor do they describe how the frequency or mean value of the trait changes over time. Population genetics models, on the other hand, are built to measure the dynamical process of evolution via the combined effects of natural selection and segregation, recombination, mutation, genetic drift, and other evolutionary forces \cite{Crow:Kimura:1970,Ewens:2004}. A natural question then is how the long-run behavior of a population genetic model (e.g., its equilibrium genotypes and induced phenotypes) compares to the equilibrium phenotypes obtained from an analogous ESS model. This question was tackled by mathematical population biologists beginning in the late 1970s and through 1980s and 1990s \cite[e..g,][]{Taylor:Jonker:1978,Hofbauer:Schuster:1979,Zeeman:1980,Eshel:1982,Hofbauer:Schuster:1982,Lessard:1984,Cressman:1988,Cressman:Hines:1984,Cressman:Hofbauer:1996,Hammerstein:1996,Weissing:1996,Eshel:1996,Eshel:Feldman:1984}. Early work in the simplest haploid asexual model with discrete strategies showed that the ESSs of obtained from condition \eqref{eq:ess:def} must be stable equilibria of haploid population genetic model \cite{Taylor:Jonker:1978,Hofbauer:Schuster:1979,Zeeman:1980}. This result isn't particularly surprising since natural selection is the only evolutionary force acting in the haploid asexual model. However, this idea was extended further by Ilan Eshel and Sabin Lessard

>>>>

Notably, the conditions in \eqref{eq:ess:def} do not reference the genetic basis of the trait $X$ nor do they describe how the frequency or mean value of the trait changes over time. Population genetics models, on the other hand, are built to measure the dynamical process of evolution via the combined effects of natural selection and segregation, recombination, mutation, genetic drift, and other evolutionary forces \cite{Crow:Kimura:1970,Ewens:2004}. A natural question then is how the long-run behavior of a population genetic model (e.g., its equilibrium genotypes and induced phenotypes) compares to the equilibrium phenotypes obtained from an analogous ESS model. This question was tackled by mathematical population biologists beginning in the late 1970s and through 1980s and 1990s \cite[e..g,][]{Taylor:Jonker:1978,Hofbauer:Schuster:1979,Zeeman:1980,Eshel:1982,Hofbauer:Schuster:1982,Lessard:1984,Cressman:1988,Cressman:Hines:1984,Cressman:Hofbauer:1996,Hammerstein:1996,Weissing:1996,Eshel:1996,Eshel:Feldman:1984}. Early work in the simplest haploid asexual population genetic model with discrete strategies showed that the ESSs of obtained from condition \eqref{eq:ess:def} must be stable equilibria of the haploid asexual model \cite{Taylor:Jonker:1978,Hofbauer:Schuster:1979,Zeeman:1980}. Specifically, suppose there are two possible phenotypes $1$ and $2$ with the payoff matrix $A = {a_{xy}}$ where $a_{xy} = \w(x,y)$ is the payoff or increase to fitness of an individual of type $x$ interacting with an individual of type $y$. Then, applying the ESS condition \eqref{eq:ess:def} to the payoff matrix $A$ gives us the stable equilibria of the haploid population genetic model. This result isn't particularly surprising since natural selection is the only evolutionary force acting in the haploid asexual model. However, this correspondence was extended to a diploid sexual population with two possible phenotypes and a single genetic locus where dominance is possible between homologous alleles and Mendelian segregation breaks up diploid genotypes every generation \cite{Eshel:1982,Lessard:1984}. Specifically, suppose there are $k$ possible alleles $A_{1},\ldots,A_{k}$ where diploid genotype $A_{i} A_{j}$ chooses phenotype $1$ with probability $v_{ij}$ and phenotype $2$ with probability $1-v_{ij}$. Eshel and Lessard show  \cite{Eshel:1982,Lessard:1984} that
